---
title: "Challenge-6"
author: "Heng Javier"
date: "20/09/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions
<br>

#### Question-1: Countdown Blastoff (While Loop)

Create a program that simulates a rocket launch countdown using a while loop. Start from 10 and countdown to "Blastoff!" with a one-second delay between each countdown number. Print a message when the rocket launches.

**Hint:** You may want to use `cat` command to print the countdown and `Sys.sleep` for incorporating the delay

**Solutions:**

```{r, eval=T,echo=TRUE}
# Enter code here
count <- 10
while (count > 0) {
        cat(count)
        cat("...")
        Sys.sleep(1)
        count = count - 1 
        }    
  cat("Blastoff!")
 


```
```    
```
<br>

#### Question-2: Word Reverser (for Loop)

Develop a program that takes a user-entered word and uses a while loop to print the word's characters in reverse order. For example, if the user enters "hello," the program should print "olleh."

**Hint:** You may want to use `substr` command to access each character of the input word, and `paste` command to join the reversed letters one at a time

**Solutions:** 

```{r, eval=T,echo=TRUE}
# Enter code here
x <- c("hello")
y <-nchar(x)
substr(x,1,y)
a<-c()
while (y>0) {
  a <- c(a, substr(x, y, y))
  y=y-1
}
paste (a, sep = " ", collapse = "", recycle0 = FALSE)

```
<br>
