---
title: "Challenge-3"
author: "Heng Javier"
date: "`30/08/2023`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>

## I. Questions

#### Question 1: Emoji Expressions

Imagine you're analyzing social media posts for sentiment analysis. If you were to create a variable named "postSentiment" to store the sentiment of a post using emojis (for positive, for neutral, for negative), what data type would you assign to this variable? Why? (_narrative type question, no code required_)

**Solution:**  Character data type. The emojis reflect ordinal data since they reflect natural ordering with no exact numerically measured distance between each of the emojis. Using the character data type, we can get a general outline of how the majority of social media users feel towards a particular social media post, or the general sentiment towards all social media posts in general. 


#### Question 2: Hashtag Havoc

In a study on trending hashtags, you want to store the list of hashtags associated with a post. What data type would you choose for the variable "postHashtags"? How might this data type help you analyze and categorize the hashtags later? (_narrative type question, no code required_)

**Solution:**  Character data type, as it is a nominal variable. To find out what the most popular hashtags are, we can just create a list and see the number of each character string that come up. Words that come up most frequently will hence be the most popular hashtags. 


#### Question 3: Time Traveler's Log
  
You're examining the timing of user interactions on a website. Would you use a numeric or non-numeric data type to represent the timestamp of each interaction? Explain your choice (_narrative type question, no code required_)

**Solution:** Numeric. If we use numeric 24 hour time, we can store the data from 0000 to 2359 to get a more specific view of the data, as compared to using bigger categories such as morning, evening and night. However, we have to be careful when taking the means of the time data as the seconds only count up to 59; a timing like 2365 does not exist. Numeric data will help us be able to compare the times from early in the morning to late at night, as bigger numbers will show before midnight and smaller numbers show after. 


#### Question 4: Event Elegance

You're managing an event database that includes the date and time of each session. What data type(s) would you use to represent the session date and time? (_narrative type question, no code required_)

**Solution:**  For the date, I would use character data type. For time, I would use numerical data type.

#### Question 5: Nominee Nominations
You're analyzing nominations for an online award. Each participant can nominate multiple candidates. What data type would be suitable for storing the list of nominated candidates for each participant?  (_narrative type question, no code required_)

**Solution:**  I would use the logical data type, since nominating a participant is a binary variable (yes or no).

#### Question 6: Communication Channels
In a survey about preferred communication channels, respondents choose from options like "email," "phone," or "social media." What data type would you assign to the variable "preferredChannel"? (_narrative type question, no code required_)

**Solution:** I would assign a character data type, since I can store responses as character strings and analyse them later. 

#### Question 7: Colorful Commentary

In a design feedback survey, participants are asked to describe their feelings about a website using color names (e.g., "warm red," "cool blue"). What data type would you choose for the variable "feedbackColor"? (_narrative type question, no code required_)

**Solution:** I would use a character data type.  

#### Question 8: Variable Exploration

Imagine you're conducting a study on social media usage. Identify three variables related to this study, and specify their data types in R. Classify each variable as either numeric or non-numeric.

**Solution:** Time spent on social media daily: integer, Numerical 
Types of social media platform used: character, Non-numeric
Purpose of using Social Media: character, Non-numeric

#### Question 9: Vector Variety

Create a numeric vector named "ages" containing the ages of five people: 25, 30, 22, 28, and 33. Print the vector.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
age <- c(25L,30L,22L,28L,33L)
age
```

#### Question 10: List Logic

Construct a list named "student_info" that contains the following elements:

  - A character vector of student names: "Alice," "Bob," "Catherine"

  - A numeric vector of their respective scores: 85, 92, 78

  - A logical vector indicating if they passed the exam: TRUE, TRUE, FALSE

Print the list.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
student_info <- list(Names=c("Alice", "Bob", "Catherine"), Score=c(85,92,78), Pass=c(T,T,F))
student_info
#OR
student_info <- list(Alice= c(Score=85,Pass=T), Bob= c(Score=92,Pass=T), Catherine = c(Score=78,Pass=F))
student_info
```

#### Question 11: Type Tracking

You have a vector "data" containing the values 10, 15.5, "20", and TRUE. Determine the data types of each element using the typeof() function.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
x<-c(10,15.5,'20',TRUE)
typeof(c(10))
typeof(c(15.5))
typeof(c('20'))
typeof(c(TRUE))
```

#### Question 12: Coercion Chronicles

You have a numeric vector "prices" with values 20.5, 15, and "25". Use explicit coercion to convert the last element to a numeric data type. Print the updated vector.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
prices<-c(20.5,15,"25")
prices<-as.numeric(prices)
prices
```

#### Question 13: Implicit Intuition

Combine the numeric vector c(5, 10, 15) with the character vector c("apple", "banana", "cherry"). What happens to the data types of the combined vector? Explain the concept of implicit coercion.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
y<-c(5,10,15)
y<-c(y,'apple','banana','cherry')
typeof(y)
#Implicit coercion refers to when numerical data is automatically converted into character data in R. 
```

#### Question 14: Coercion Challenges

You have a vector "numbers" with values 7, 12.5, and "15.7". Calculate the sum of these numbers. Will R automatically handle the data type conversion? If not, how would you handle it?

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
z<-c(7,12.5,'15.7')
#sum(z)
#R cannot convert the character  '15.7'. As such, we have to either  manually convert it ourselves or use explicit coercion.
z<-as.double(z)
sum(z)
typeof(sum(z))
```

#### Question 15: Coercion Consequences

Suppose you want to calculate the average of a vector "grades" with values 85, 90.5, and "75.2". If you directly calculate the mean using the mean() function, what result do you expect? How might you ensure accurate calculation?

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
grades <- c(85,90.5,'75.2')
mean(grades)
#NA result since '75.2' is a character data type. To ensure accurate calculation, either convert it manually by removing the quotation marks or use explicit coercion.
grades<- as.double(grades)
mean(grades)
```

#### Question 16: Data Diversity in Lists

Create a list named "mixed_data" with the following components:

  - A numeric vector: 10, 20, 30
  
  - A character vector: "red", "green", "blue"

  - A logical vector: TRUE, FALSE, TRUE
  
Calculate the mean of the numeric vector within the list.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
mixed_data <- list(Numeric= c(10,20,30),colour= c('red','green','blue') ,Logical=c( T,F,T))
mixed_data
mean(mixed_data$Numeric)
```

#### Question 17: List Logic Follow-up

Using the "student_info" list from Question 10, extract and print the score of the student named "Bob."

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
student_info <- list(Alice= c(Score=85,Pass=T), Bob= c(Score=92,Pass=T), Catherine = c(Score=78,Pass=F))
student_info
student_info$Bob["Score"]
```

#### Question 18: Dynamic Access

Create a numeric vector values with random values. Write R code to dynamically access and print the last element of the vector, regardless of its length.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
b<-c(1:80)
b<-c(1,3,6,7,2)
b[length(b)]
#The bottom-most code should give the last element for both of the aforementioned vectors. 

```

#### Question 19: Multiple Matches

You have a character vector words <- c("apple", "banana", "cherry", "apple"). Write R code to find and print the indices of all occurrences of the word "apple."

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
words <- c("apple", "banana", "cherry", "apple")
print(words=="apple")
which(words=="apple")
```

#### Question 20: Conditional Capture

Assume you have a vector ages containing the ages of individuals. Write R code to extract and print the ages of individuals who are older than 30.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
age<-c(10,20,30,40,50,60)
age[age>30]
```

#### Question 21: Extract Every Nth

Given a numeric vector sequence <- 1:20, write R code to extract and print every third element of the vector.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
q<-1:20
which(q%%3==0)
#or
q <- q[seq(0,length(q),3)]
q
```

#### Question 22: Range Retrieval

Create a numeric vector numbers with values from 1 to 10. Write R code to extract and print the values between the fourth and eighth elements.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
d<-(1:10)
d[5:7]
```

#### Question 23: Missing Matters

Suppose you have a numeric vector data <- c(10, NA, 15, 20). Write R code to check if the second element of the vector is missing (NA).

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
data <- c(10, NA, 15, 20)
is.na(data[2])
#NA is unknown and undefined by r, so == doesn't work. 
```

#### Question 24: Temperature Extremes

Assume you have a numeric vector temperatures with daily temperatures. Create a logical vector hot_days that flags days with temperatures above 90 degrees Fahrenheit. Print the total number of hot days.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
temperatures<-c(70,80,90,100,110,120)
hot_days <- temperatures>90
hot_days
sum(temperatures>90)
#or
length(which(hot_days==TRUE))
```

#### Question 25: String Selection

Given a character vector fruits containing fruit names, create a logical vector long_names that identifies fruits with names longer than 6 characters. Print the long fruit names.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
fruits <- c("apple","banana","cherry","pineapple","tomato")
long_names <- as.logical(nchar(fruits)>6)
long_names
fruits[long_names]  
#or
fruits[long_names==TRUE]
```

#### Question 26: Data Divisibility

Given a numeric vector numbers, create a logical vector divisible_by_5 to indicate numbers that are divisible by 5. Print the numbers that satisfy this condition.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
v<-c(1,4,5,10,16,20)
divisible_by_5 <- as.logical(v%%5==0)
divisible_by_5
v[divisible_by_5]
#or
v[v%%5==0]
#or
v[divisible_by_5==TRUE]
```

#### Question 27: Bigger or Smaller?

You have two numeric vectors vector1 and vector2. Create a logical vector comparison to indicate whether each element in vector1 is greater than the corresponding element in vector2. Print the comparison results.

**Solution:** 

```{r, eval=T,echo=TRUE}
# Enter code here
vector1<-c(2,3,4,5,6)
vector2<-c(10,8,6,4,2)
vector3<-vector1>vector2
vector3
#Doesn't work if string length of vector1 =/= string length of vector 2!
```

