---
title: "Week 9 Project Submission"
author: "Heng Javier"
date: "2023-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Week 9 Questions:

# (1) What is the topic that you have finalized? (Answer in 1 or 2 sentences).
I will be doing an analysis of Taylor Swift's music/albums on Spotify.Specifically, I will be focusing on aspects or her music such as tempo, valence and time signature.

# (2) What are the data sources that you have curated so far? (Answer 1 or 2 sentences).
I will be using the Taylor Swift dataset from Tidy Tuesday.

Week 10 Questions:

# (1) What is the question that you are going to answer? (Answer: One sentence that ends with a question mark that could act like the title of your data story),
What is Taylor Swift's music like? 
# (2) Why is this an important question? (Answer: 3 sentences, each of which has some evidence, e.g., “According to the United Nations...” to justify why the question you have chosen is important),
In August 2023, Taylor Swift was the first female artist to reach 100 million monthly listeners. 
Taylor Swift has a $740 million net worth as of June 2023 (according to Forbes).
Analysing the stats of her music will help other musicians to figure out the ingredients in her music to cook up hits just like her.
# (3) Which rows and columns of the dataset will be used to answer this question? (Answer: Actual names of the variables in the dataset that you plan to use).
I will use track_name, danceability, energy, key, 

```{r,echo=TRUE,eval=TRUE}
library(tidyverse)
install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2023-10-17')

taylor_album_songs <- tuesdata$taylor_album_songs
taylor_all_songs <- tuesdata$taylor_all_songs
taylor_albums <- tuesdata$taylor_albums

taylor_album_songs
```
