---
pdf_document: default
author: "Heng Javier"
date: "`29/08/2023`"
output: pdf_document
title: "Week-3: Code-along"
---
x
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# I. Code to edit and execute

**To be submitted on canvas before attending the tutorial**


### Loading packages

```{r, eval=T,echo=TRUE}
# Load package tidyverse
library(tidyverse)
```

### Assigning values to variables

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Example a.: execute this example
x <- 'A'
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example b and execute it
x <- 'Apple'
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example c and execute it
x <- FALSE
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example d and execute it
x<-5L
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example e and execute it
x<-5
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example f and execute it
x<-1i
```

### Checking the type of variables

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Example a.: execute this example
x <- 'A'
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example b and execute it
x <- 'Apple'
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example c and execute it
x <- FALSE
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example d and execute it
x<-5L
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example e and execute it
x<-5
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Complete the code for Example f and execute it
x<-1i
typeof(x)
```

### Need for data types

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# import the cat-lovers data from the csv file you downloaded from canvas
df <- read.csv("cat-lovers.csv")

```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Compute the mean of the number of cats: execute this command
mean(df$number_of_cats)
```

```{r,warning=TRUE,message=FALSE,eval=FALSE,echo=TRUE}
# Get more information about the mean() command using ? operator
?mean
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Convert the variable number_of_cats using as.integer()
as.integer(df$number_of_cats)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Display the elements of the column number_of_cats 
df$number_of_cats
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Display the elements of the column number_of_cats after converting it using as.numeric()
as.numeric(df$number_of_cats)
```

### Create an empty vector

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Empty vector
x <- c()
# Type of the empty vector
typeof(x)
```

### Create vectors of type logical

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 1
x<-vector("logical",length=5)
# Display the contents of x
print(x)
# Display the type of x
print(typeof(x))
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 2
x<-logical(5)
# Display the contents of x
print(x)
# Display the type of x
print(typeof(x))
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 3
x<-c(TRUE,FALSE,TRUE,FALSE,TRUE)
# Display the contents of x
print(x)
# Display the type of x
print(typeof(x))
```

### Create vectors of type character

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 1
x<-vector("character",length=5)
# Display the contents of x
print(x)
# Display the type of x
print(typeof(x))
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 2
x <- character(5)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)

```

```{r,warning=TRUE,message=FALSE,eval=TRUE,echo=TRUE}
# Method 3
x<-c('a','b','c','d','e')
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

### Create vectors of type integer

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 1
x<-vector("integer",length=5)
# Display the contents of x
print(x)
# Display the type of x
print(typeof(x))
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 2
x<-integer(5)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 3
x<-c(1L,2L,3L,4L,5L)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)

```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 4
x <- seq(from=1L,to=5L,by=1L)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 5
 x<-1:5

# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

### Create vectors of type double

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 1
x<-vector("double",length=5)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 2
x<-double(5)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

```{r,warning=TRUE,message=FALSE,eval=T,echo=TRUE}
# Method 3
x<-c(1,1.5,2,2.5,3)
# Display the contents of x
print(x)
# Display the type of x
typeof(x)
```

### Implicit coercion

#### Example 1 (Conversion of types from double to character)

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x<-c(1.8)
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Add a character to the vector
x<-c(x,'a')
print(x)
# Check the type of x
typeof(x)
```

#### Example 2 (Conversion of types from logical to double)

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x<-c(T)
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Add a number to the vector
x<-c(x,2)
print(x)
# Check the type of x
typeof(x)
```

#### Example 3 (Conversion of types from character to character)

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x<-c(x,TRUE)
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Add a logical value to the vector
x<-(F)
# Check the type of x
typeof(x)
```


#### Example 4 (Conversion of types from integer to double)

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x <-c(1L)
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Add a number to the vector
x<-c(x,2)
# Check the type of x
typeof(x)
```

### Explicit coercion

#### Example 1(From integer to character)

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x<-c(1L)
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Convert the vector to type character
x<-as.character(x)
print(x)
# Check the type of x
typeof(x)
```

#### Example 2 (from character to double)

```{r,warning=TRUE,message=TRUE,eval=FALSE,echo=TRUE}
# Create a vector
x<-c('A')
# Check the type of x
typeof(x)
```

```{r,warning=TRUE,message=TRUE,eval=FALSE,echo=TRUE}
# Convert the vector to type double
x<-as.numeric(x)
# Check the type of x
typeof(x)
```

### Accessing elements of the vector

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Create a vector
x <- c(1,10,9,8,1,3,5)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Access one element with index 3
x[3]
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Access elements with consecutive indices, 2 to 4: 2,3,4
x[c(2:4)]
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Access elements with non-consecutive indices, 1,3,5
x[c(1,3,5)]
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Access elements using logical vector
x[c(TRUE,FALSE,FALSE,TRUE,FALSE,FALSE,TRUE)]
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Access elements using the conditional operator <
x[x>0]
x[x<9]
```

### Examining vectors

```{r,warning=TRUE,message=TRUE,eval=TRUE,echo=TRUE}
# Display the length of the vector
length(x)
# Display the type of the vector
typeof(x)
# Display the structure of the vector
str(x)
```

### Lists

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Initialise a named list
my_pie = list(type="key lime", diameter=7, is.vegetarian=TRUE)
# display the list
my_pie
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Print the names of the list
names(my_pie)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Retrieve the element named type
my_pie$type
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Retrieve a truncated list
my_pie['type']
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Retrieve the element named type
my_pie[['type']]
```

#### Exploring data-sets

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Install package
install.packages("openintro", repos="http://cran.us.r-project.org")
# Load the package
library(openintro)
# Load package
library(tidyverse)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Catch a glimpse of the data-set: see how the rows are stacked one below another
glimpse(loans_full_schema)
```


```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Selecting numeric variables
loans <- loans_full_schema %>% # <-- pipe operator
  select(paid_total, term, interest_rate,
         annual_income,paid_late_fees,debt_to_income)
# View the columns stacked one below another
glimpse(loans)
```

```{r,warning=TRUE,message=TRUE,eval=T,echo=TRUE}
# Selecting categoric variables
loans <- loans_full_schema %>% 
  select( ) # type the chosen columns as in the lecture slide
# View the columns stacked one below another
glimpse(loans)
```